Simulation Study of Angular Resolution of a New Electromagnetic Sampling Calorimeter

### 2022-05-30 paper.tex
updates

\begin{abstract}
We report simulation results on the angular resolution of electromagnetic (EM) sampling calorimeters with photons being in the range of 100~MeV to 2~GeV. A simulation model of the EM calorimeter consists of alternating layers of a 1-mm-thick Pb plate and a 5-mm-thick plastic scintillator plate. The scintillator plates are alternatively segmented in horizontal and vertical strips. In this study, we obtained energy deposits in individual strips using Geant4 simulations and reconstructed incident photon angles using a XGBoost model with boosted decision trees. Angular resolutions are studied in terms of the strip width. The 15-mm-wide strips provide the best angular resolution of 1.23~$\pm$~0.01$^{\circ}$. Incident photon directions can be well reconstructed in the front part (5$X_{0}$) of the EM calorimeter with the acceptable angular resolution of 1.33~$\pm$~0.01$^{\circ}$. The angular resolution is mildly dependent on the incidence angle, and varies significantly with the incident photon energy. The energy dependence can be expressed as $\sigma_{\theta}=$~0.43~$\oplus p_{1} /\sqrt{E_{\gamma}}$, where $p_{1}$ gradually decreases with the incidence angle.

@@

This paper describes a machine learning approach based on XGBoost (XGB) model~\cite{xgboost:2016} with boosted decision trees in order to reconstruct the direction of incident photons entering a sampling calorimeter with alternating layers of a Pb plate and plastic scintillator strips in horizontal and vertical directions. We generated EM showers in a sampling calorimeter using Geant4 simulations. We report on the detector configuration providing the acceptable angular resolution, and report on the angular resolution of the EM sampling calorimeter in terms of the incident angle and energy.

@@

Figure~\ref{fig:angle_10degree} represents a distribution of reconstructed incidence angles $\theta$ for 1-GeV photons generated at $\theta=$~10$^{\circ}$. The distribution is fitted with the Gaussian function and the Generalized Gaussian (GG) function. We tested two functions, Gaussian and GG to describe the reconstructed angle distribution. Two functions were multiplied by $\sin{x}$ to correct solid angle acceptance. The GG function, also known as the generalized error distribution, is expressed as
\begin{eqnarray} 
f(x; \mu, \alpha, \beta) = \frac{\beta}{2 \alpha \Gamma(1/\beta)}e^{-(|x-\mu|/\alpha)^\beta},
\label{eqn:gg}
\end{eqnarray}
where $\mu$ denotes a mean value. Parameters $\alpha$ and $\beta$ determine the scale and shape of the distribution, respectively. Variance of the GG function is given by $\sigma^2 \equiv \alpha^2 \Gamma(3/\beta) / \Gamma(1/\beta)$. The angular resolution of the incidence angle reconstruction is defined as $\sigma$.

@@

Furthermore, incidence angles are reconstructed with the front layers of $X_{0}$ in which 99\% of incident photons generate EM showers. Figure~\ref{fig:angle_reco_layer} (a) shows reconstructed angles using front 24 layers of the detector, which correspond to 4.6$X_{0}$ for 1-GeV photons. A fraction of photons failed to be reconstructed due to the lack of active layers. The failed events are represented as a delta function near 0, and such events are removed by requiring the total energy deposit to be larger than 10~MeV, 1\% of incident energy. The angular resolution with the front layer is estimated to be 1.33~$\pm$~0.01$^{\circ}$ while possessing inefficiency estimated to be 5.8\%.

>>>>

\begin{abstract}
We report simulation results on the angular resolution of electromagnetic (EM) sampling calorimeters with photons being in the range of 100~MeV to 2~GeV. A simulation model of the EM calorimeter consists of alternating layers of a 1-mm-thick Pb plate and a 5-mm-thick plastic scintillator plate. The scintillator plates are alternatively segmented in horizontal and vertical strips. In this study, we obtained energy deposits in individual strips using Geant4 simulations and reconstructed incident photon angles using a XGBoost model with boosted decision trees. Angular resolutions are studied in terms of the strip width. The 15-mm-wide strips provide the best angular resolution of 1.23~$\pm$~0.01$^{\circ}$. Incident photon directions can be well reconstructed in the front part (4.6$X_{0}$) of the EM calorimeter with the acceptable angular resolution of 1.33~$\pm$~0.01$^{\circ}$. The angular resolution is mildly dependent on the incidence angle, and varies significantly with the incident photon energy. The energy dependence can be expressed as $\sigma_{\theta}=$~0.43~$\oplus p_{1} /\sqrt{E_{\gamma}}$, where $p_{1}$ gradually decreases with the incidence angle.

@@

This paper describes a machine learning approach based on XGBoost (XGB) model~\cite{xgboost:2016} with boosted decision trees in order to reconstruct the direction of incident photons entering a sampling calorimeter with alternating layers of a Pb plate and plastic scintillator strips in horizontal and vertical directions. We generated EM showers in a sampling calorimeter using Geant4 simulations. We report on the detector configuration providing the acceptable angular resolution, and report on the angular resolution of the EM sampling calorimeter in terms of the incidence angle and the incident energy.

@@

Figure~\ref{fig:angle_10degree} represents a distribution of the relative incidence angle ($\Delta\theta$) for 1-GeV photons generated at $\theta=$~10$^{\circ}$. $\Delta\theta$ denotes how much distant the reconstructed incidence direction is from the true incidence direction radially. The central 98\% part of the distribution is fitted with the Gaussian function and the Generalized Gaussian (GG) function. We tested two functions, Gaussian and GG to describe the reconstructed angle distribution. The GG function, also known as the generalized error distribution, is expressed as
\begin{eqnarray} 
f(x; \mu, \alpha, \beta) = \frac{\beta}{2 \alpha \Gamma(1/\beta)}e^{-(|x-\mu|/\alpha)^\beta},
\label{eqn:gg}
\end{eqnarray}
where $\mu$ denotes a mean value. Parameters $\alpha$ and $\beta$ determine the scale and shape of the distribution, respectively. Variance of the GG function is given by $\sigma^2 \equiv \alpha^2 \Gamma(3/\beta) / \Gamma(1/\beta)$. The angular resolution of the incidence angle reconstruction is defined as $\sigma$.

@@

Furthermore, incidence angles are reconstructed with the front layers of $4.6X_{0}$ in which 99\% of incident photons generate EM showers. Figure~\ref{fig:angle_reco_layer} (a) shows reconstructed angles using front 24 layers of the detector, which correspond to 4.6$X_{0}$ for 1-GeV photons. A fraction of photons failed to be reconstructed due to the lack of active layers. The failed events are represented as a delta function near 0, and such events are removed by requiring the total energy deposit to be larger than 10~MeV, 1\% of the incident energy. The angular resolution with the front layer is estimated to be 1.33~$\pm$~0.01$^{\circ}$ while possessing inefficiency estimated to be 5.8\%.
